Type theories as quotient inductive-recursive types

### 2025-02-11 tex/main.tex
Use single quotes instead of double quotes

We write $x \jdeq y$ for strict equality, $x =^{A} y$ for the identity type stating that elements $x : A$ and $y : A$ of a type $A$ are equal, and $\dpath{P}{p}{t}{u}$ for the ``dependent'' identity type where $P : A \to \Set$, $t : P\,x$, $u : P\,y$ and $p : x =^A y$. One possible realisation of this type is as $t =^{P}_{p} u \defeq \transfib{P}{p}{t} =^{P\,y} u$, where $\transfibf{P} : (p : x =^A y) \to P\,x \to P\,y$ is our notation for the operation which allows is to transport values between type families applied to elements that are equal up to the identity type.
%
%Finally, we write $x \simeq y$ for the \emph{heterogeneous} identity type, where we allow $x : A$ and $y : B$ to have potentially different types.
%As shown by McBride~\cite{McBride1999}, the heterogeneous identity type is equivalent to the ordinary identity type equipped with the principle of uniqueness of identity proofs.

We have formally implemented our definitions and proofs in Agda, with the following options: \texttt{-{}-with-K}, \texttt{-{}-exact-split}, \texttt{-{}-rewriting}, and \texttt{-{}-local-confluence-check}. We simulate quotient inductive types using postulated equations to introduce equality constructors~\cite{Licata2011}.
Some of our theorem statements are meta-theoretical, and thus cannot be checked by Agda. Others can only be checked  partially, because of postulated quotient inductive types. We have marked such ``informal'' proofs with the symbol \resizebox*{!}{\baselineskip}{\textdbend}.
\LT{Single quote or double quote in British English?}

@@

In contrast, an inductive-recursive type is defined inductively \emph{simultaneously} with a recursively defined function out of it. The prototypical example is Martin-L\"of's universe \`a la Tarski, where a type $U$ of codes for small types is defined simultaneously with a ``decoding'' function $T : U \to \Set$ which maps each code to the type it represents. Crucially, the code for dependent function types
\[
  \pi : (a : U) \to (T(a) \to U) \to U
\]
with $T(\pi(a, b)) = (x : T(a)) \to T(b(x))$ only makes sense because we already have access to the decoding function $T$.
%
Inductive-recursive types (IRT) were introduced by Dybjer~\cite{Dybjer2000}, and then extensively studied by Dybjer and Setzer~\cite{Dybjer1999,Dybjer2003}.
The general schema allows one to define a type $A$ inductively, together with a recursive function $F : A \to D$, for some fixed (possibly large) type $D$.

>>>>

We write $x \jdeq y$ for strict equality, $x =^{A} y$ for the identity type stating that elements $x : A$ and $y : A$ of a type $A$ are equal, and $\dpath{P}{p}{t}{u}$ for the `dependent' identity type where $P : A \to \Set$, $t : P\,x$, $u : P\,y$ and $p : x =^A y$. One possible realisation of this type is as $t =^{P}_{p} u \defeq \transfib{P}{p}{t} =^{P\,y} u$, where $\transfibf{P} : (p : x =^A y) \to P\,x \to P\,y$ is our notation for the operation which allows is to transport values between type families applied to elements that are equal up to the identity type.
%
%Finally, we write $x \simeq y$ for the \emph{heterogeneous} identity type, where we allow $x : A$ and $y : B$ to have potentially different types.
%As shown by McBride~\cite{McBride1999}, the heterogeneous identity type is equivalent to the ordinary identity type equipped with the principle of uniqueness of identity proofs.

We have formally implemented our definitions and proofs in Agda, with the following options: \texttt{-{}-with-K}, \texttt{-{}-exact-split}, \texttt{-{}-rewriting}, and \texttt{-{}-local-confluence-check}. We simulate quotient inductive types using postulated equations to introduce equality constructors~\cite{Licata2011}.
Some of our theorem statements are meta-theoretical, and thus cannot be checked by Agda. Others can only be checked  partially, because of postulated quotient inductive types. We have marked such `informal' proofs with the symbol \resizebox*{!}{\baselineskip}{\textdbend}.
\LT{Single quote or double quote in British English?}

@@

In contrast, an inductive-recursive type is defined inductively \emph{simultaneously} with a recursively defined function out of it. The prototypical example is Martin-L\"of's universe \`a la Tarski, where a type $U$ of codes for small types is defined simultaneously with a `decoding' function $T : U \to \Set$ which maps each code to the type it represents. Crucially, the code for dependent function types
\[
  \pi : (a : U) \to (T(a) \to U) \to U
\]
with $T(\pi(a, b)) = (x : T(a)) \to T(b(x))$ only makes sense because we already have access to the decoding function $T$.
%
Inductive-recursive types (IRT) were introduced by Dybjer~\cite{Dybjer2000}, and then extensively studied by Dybjer and Setzer~\cite{Dybjer1999,Dybjer2003}.
The general schema allows one to define a type $A$ inductively, together with a recursive function $F : A \to D$, for some fixed (possibly large) type $D$.
